### 2013-02-28 doc/report/intro.tex
added docs and LICENSE

\section{Introduction}
\label{sec:introduction}

So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: