### 2013-04-05 doc/report/intro.tex
some old check in stuff, just cleaned it

\section{Introduction}
\label{sec:introduction}

Traditional ways of interacting with applications on phones or tablets are browsing the available applications and click on them. You may have the freedom to set shortcuts on home screen for easy access. While this is unavoidable in cyber world, for interaction with physical devices, there is a more natural way to select by exploiting spatial information - proximity and attention.

So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: