### 2013-05-03 doc/report/conclusion.tex
adding files to the report

\section{Conclusion}
\label{sec:conclusion}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: