### 2013-05-03 doc/report/design.tex

\section{Design}
\label{sec:design}

Talks about the system design space.
What might be available to achieve such a function.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: