### 2013-05-03 doc/report/discussion.tex

\section{Discussion}
\label{sec:discussion}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: