### 2013-05-03 doc/report/evaluation.tex

\section{Evaluation}
\label{sec:evaluation}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: