### 2013-05-03 doc/report/implementation.tex

\section{Implementation}
\label{sec:implementation}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: