### 2013-05-03 doc/report/main.tex

\documentclass{sig-alternate}

\usepackage{times}
\usepackage{graphics}

\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{tabularx}
\usepackage{color}
\usepackage{xspace}
\usepackage{hyperref}    % Creates hyperlinks from ref/cite
\hypersetup{pdfstartview=FitH}
\usepackage{graphicx}    % For importing graphics
\usepackage{url}         %

\hypersetup{%
pdftitle={Project Glass 2.0}, pdfauthor={Ben Zhang, Sean Chen}, pdfkeywords={HCI}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
breaklinks=true,}

\renewcommand{\arraystretch}{1.2} % Space out rows in tables

\newcommand{\ml}[1]{{\color{green} {\it ML: #1}}}

% No space between bibliography items:
\let\oldthebibliography=\thebibliography
  \let\endoldthebibliography=\endthebibliography
  \renewenvironment{thebibliography}[1]{%
    \begin{oldthebibliography}{#1}%
      \setlength{\parskip}{0ex}%
      \setlength{\itemsep}{0ex}%
  }%
  {%
    \end{oldthebibliography}%
  }

%\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy

\begin{document}

% use this command to override the default ACM copyright statement
% (e.g. for preprints). Remove for camera ready copy.
%\toappear{Submitted for review to IPSN 2012.}
% \conferenceinfo{ConfName} {Date, Location}
% \CopyrightYear{Year} 
% \crdata{978-1-4503-1227-1/12/04} 
% \clubpenalty=10000 
% \widowpenalty = 10000

% to make various LaTeX processors do the right thing with page size
\special{papersize=8.5in,11in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\title{Project Glass 2.0 -- Interaction with Physical Devices through Attention}

\author{
{Ben Zhang, Sean Chen}\\
\affaddr{University of California, Berkeley}\\
%\affaddr{}\\
\email{benzh@eecs.berkeley.edu, sean.yhc@berkeley.edu}
}

\maketitle

\begin{abstract}
With the increasing interest of assigning intelligence to everyday object, many approaches of interacting with them fall back to smartphone based applications. We argue about the inconvenience, overhead, and distraction in such an interaction paradigm, and propose a more natural solution. A glass-based controller which enables user to express their interest in interaction directly simplifies the process of selection in a smart-home or smart-office senario. We present our thoughts, design and evaluation of such a system in this paper to motivate researchers in considering ``interaction in attention''. In the meantime, we noticed that Google Glass might serve as the perfect enabling technology, though in our prototype we've adopted simple infrared approach for proof-of-concept.
\end{abstract}

%\category{C.2.1}{Computer-Communication Networks}{Network Architecture and Design}[Wireless communication]
%\keywords{Collection, CTP, Sensor Network, Routing}

% \category{B.0}{Hardware}{General}
% \category{B.4}{Hardware}{Input/Output \& Data Communication} 
% \category{H.4.m}{Information Systems Applications}{Miscellaneous}

% \terms{Design, Experimentation, Measurement} 
\keywords{HCI, Interaction, Attention, Infrared, Glass}

\input{intro.tex}
\input{related.tex}
\input{design.tex}
\input{implementation.tex}
\input{evaluation.tex}
\input{discussion.tex}
\input{conclusion.tex}

% \section{Acknowledgments}

\bibliographystyle{abbrv}
\bibliography{reference}

\end{document}