### 2013-05-05 doc/report/conclusion.tex
added some texts to the report, and a glass FSM figure

\section{Conclusion}
\label{sec:conclusion}

In this project, we explored the novel way of interacting with physical device by expressing users' attention. The prototype we have built demonstrate the possibility of identifying user's attention and control various devices.

We will leave X, Y, Z to future work.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: