### 2013-05-05 doc/report/discussion.tex

\section{Discussion}
\label{sec:discussion}

\textcolor{red}{Discuss the problems we have encountered in this section.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: