### 2013-05-05 doc/report/evaluation.tex

\section{Evaluation}
\label{sec:evaluation}

\textcolor{red}{Will fill this evaluation when we get to it.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: