### 2013-05-06 doc/report/main.tex
changed to sigchi style “

\documentclass{sigchi}

\usepackage{times}
\usepackage{graphics}

\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{tabularx}
\usepackage{color}
\usepackage{xspace}

\usepackage{graphicx}    % For importing graphics
\usepackage{url}         %

\renewcommand{\arraystretch}{1.2} % Space out rows in tables

\newcommand{\ml}[1]{{\color{green} {\it ML: #1}}}

% No space between bibliography items:
\let\oldthebibliography=\thebibliography
  \let\endoldthebibliography=\endthebibliography
  \renewenvironment{thebibliography}[1]{%
    \begin{oldthebibliography}{#1}%
      \setlength{\parskip}{0ex}%
      \setlength{\itemsep}{0ex}%
  }%
  {%
    \end{oldthebibliography}%
  }

\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy

% Use this command to override the default ACM copyright statement (e.g. for preprints). 
% Consult the conference website for the camera-ready copyright statement.
\toappear{
Permission to make digital or hard copies of all or part of this
work for personal or classroom use is granted without fee provided that 
copies are not made or distributed for profit or commercial advantage and
that copies bear this notice and the full citation on the first page. To
copy otherwise, or republish, to post on servers or to redistribute to 
lists, requires prior specific permission and/or a fee.\\
Copyright 20XX ACM X-XXXXX-XX-X/XX/XX ...\$10.00.
}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% to make various LaTeX processors do the right thing with page size
\special{papersize=8.5in,11in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={Project Glass 2.0}, 
pdfauthor={Ben Zhang, Sean Chen},
pdfkeywords={HCI},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{Project Glass 2.0 -- Interaction with Physical Devices through Attention}

\numberofauthors{2}
\author{
{Ben Zhang, Sean Chen}\\
\affaddr{University of California, Berkeley}\\
%\affaddr{}\\
\email{benzh@eecs.berkeley.edu, sean.yhc@berkeley.edu}
}

\maketitle

\begin{abstract}
With the increasing interest of assigning intelligence to everyday object, many approaches of interacting with them fall back to smartphone based applications. We argue about the inconvenience, overhead, and distraction in such an interaction paradigm, and propose a more natural solution. A glass-based controller which enables user to express their interest in interaction directly simplifies the process of selection in a smart-home or smart-office senario. We present our thoughts, design and evaluation of such a system in this paper to motivate researchers in considering ``interaction in attention''. In the meantime, we noticed that Google Glass might serve as the perfect enabling technology, though in our prototype we've adopted infrared approach for proof-of-concept. The evaluation and discussion of such system might be valuable for future work to explore on more mature hardware platform.
\end{abstract}

%\category{C.2.1}{Computer-Communication Networks}{Network Architecture and Design}[Wireless communication]
%\keywords{Collection, CTP, Sensor Network, Routing}

% \category{B.0}{Hardware}{General}
% \category{B.4}{Hardware}{Input/Output \& Data Communication} 
% \category{H.4.m}{Information Systems Applications}{Miscellaneous}

% \terms{Design, Experimentation, Measurement} 
\keywords{HCI, Interaction, Attention, Infrared, Glass}

\input{intro.tex}
\input{related.tex}
\input{design.tex}
\input{implementation.tex}
\input{evaluation.tex}
\input{discussion.tex}
\input{conclusion.tex}

% \section{Acknowledgments}

\bibliographystyle{abbrv}
\bibliography{reference}

\end{document}