### 2013-05-07 doc/report/discussion.tex
finished implementation, and start discussion

\section{Discussion}
\label{sec:discussion}

\textcolor{red}{Discuss the problems we have encountered and other random stuffs in this section.}

\subsection{Interaction in Attention}
\label{sec:inter-attent}

During our brainstorming and discussion of this idea, we have identified some tasks where user might not pay their attention to the device they are controlling. One such example is giving presentation. The attention of presenter is to interact with the audience, but this is achieved through controlling the slides and projection.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: