### 2013-05-07 doc/report/discussion.tex
added some discussion

\section{Discussion}
\label{sec:discussion}

% Discuss the problems we have encountered and other random stuffs in this section
In this section, we will touch something that is related to our project, or pops up during our thinking that might be beneficial to people who consider working on similar topics. We will first talk about the attention-based interaction and its limitation. Then specific to our project implementation, we would like to express some of our experience.

\subsection{Interaction in Attention}
\label{sec:inter-attent}

During our brainstorming and discussion of this idea, we have identified some tasks where user might not pay their attention to the device they are controlling. One such example is giving presentation. The attention of presenter is to interact with the audience, but this is achieved through controlling the slides and projection.

{\color{red} a very short brainstorming would be able to finish this part.}

\subsection{Experience learned in this project}
\label{sec:exper-learn-this}

XBee setting up.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: