### 2013-05-19 doc/report/evaluation.tex
update evaluation

\section{Evaluation}
\label{sec:evaluation}

\textcolor{red}{Will fill this evaluation when we get to it.}
Some of the ideas as to how to work on evaluation include: (A) Make a mobile (web) app that can remotely control the appliances through the gateway. Give a sequence of tasks to the tester and have him complete the tasks first on smartphone then use the Glass. (B) Observe the differences between using LED, Recon screen, and other feedbacks such as audio, and how that affects the user experience.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: