### 2013-05-19 doc/report/evaluation.tex
typos, and removed some red texts, also removed .pdf file..., compile the src when you need, don't include pdf file in it

\section{Evaluation}
\label{sec:evaluation}

%\textcolor{red}{Will fill this evaluation when we get to it.}
Constrained by the time within a semester, we haven't conducted a formal user study. Based on our discussions with the instructors, we have some initial plans as to how to work on evaluation -- (A) Make a mobile (web) app that can remotely control the appliances through the gateway. Give a sequence of tasks to the tester and have him complete the tasks first on smartphone then use the Glass. (B) Observe the differences between using LED, Recon screen, and other feedbacks such as audio, and how that affects the user experience.

We are planning to have evaluation done within the summer.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: