### 2013-09-02 doc/sigchi14/control.tex
bjoern's pass at overall narrative

\section{Example Scenarios}
\bjoern{Here we should either describe a couple of example scenarios we've built, or get some students to run through realistic control tasks based on those scenarios - e.g.: ``Dim the light to 20\% and then start the movie, turn up the volume to 80\%."}