### 2013-09-03 doc/sigchi14/conclusion.tex

\section{Conclusion and Future Work}
Our prototype does not yet have a general way of communicating device descriptions and capabilities - this means our prototype cannot yet control new, unknown devices - they first have to be modeled. We may extend the PUC work~\cite{nichols} to address this issue.

More importantly though, the main limitation is that extra hardware for IR communication is needed for the HMD and each controllable device. One potential approach would be to combine the growing availability of high-resolution indoor maps with live data from the HMD's camera to determine what a user is looking at without IR data exchange.

\bjoern{add some general summary}