### 2013-09-13 doc/sigchi14/discussion.tex

\section{Discussion}

Study-related topics:
\begin{enumerate}
\item Our technique should have a wider margin as the number of targets increases.

\end{enumerate}

Other topics:
\begin{enumerate}
\item Some fundamental differences to earlier laser-pointer based systems: with a laser pointer in hand one can very precisely point at an arbitrary object and know that one is pointing; however, that alone does not provide feedback on whether a device can actually be controlled or not. By shifting the visual feedback to the target device,

\item Battery life considerations: we still need to know when to start emitting the IR signal, otherwise, battery life will suffer, which is especially critical for wireless devices. However, requiring complicated steps to turn IR on goes counter the design goals of seamless interactions. A frame-mounted button may be an appropriate design choice.

\item Where to look? For a light, one might put a received on the light itself, or on the light switch. For volume, it might be on a speaker or on the amplifier. Someone should study preferences.
\end{enumerate}