### 2013-09-18 doc/sigchi14/abstract.tex
abstract

We introduce a novel method for selecting and controlling smart appliances in physical spaces through a head-worn computing device with near-eye display and wireless communication. We augment a commercial wearable computing device, Google Glass, with a narrow-beam IR emitter for this purpose. This configuration yields a usable beam width of 2 to 4 feet (60 to 120cm) for targeting at room scale. We  describe a disambiguation technique if infrared targeting hits multiple targets simultaneously. A target acquisition study with 15 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. We also report qualitative data from using our device to control multiple appliances in a smart home scenario.