### 2013-09-18 doc/sigchi14/conclusion.tex

\section{Conclusion}
We introduced a novel method for selecting and controlling smart appliances in physical spaces through  infrared targeting based on head orientation. While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware. We also introduced a disambiguation technique in case head orientation is not sufficient to determine a unique target. We characterized our devices performance, arguing that it is matched well to the amount of head movement people can control without strain. A target acquisition study showed that the technique is efficient; a home control scenario showed promise but also limitations when trying to control complex appliances. As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments may become increasingly important.
%\section{Acknowledgments}
%We thank our user study participants and the a