### 2013-09-18 doc/sigchi14/conclusion.tex
refinments

\section{Conclusion}
We introduced a novel method for selecting and controlling smart appliances in physical spaces through infrared targeting based on head orientation. Through our solution, we attempt to address the naming and scaling challenges faced by handheld mobile devices. The design takes advantage of the fact that visual attention can express intention. The visual feedback provided by the target appliances helps users keep their focus in the physical world. While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware. We also introduced a disambiguation technique in case head orientation is not sufficient to determine a unique target. We characterized our devices performance, arguing that it is matched well to the amount of head movement people can control without strain. A target acquisition study showed that the technique is efficient; a home control scenario showed promise but also limitations when trying to control complex appliances. As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments will become increasingly important.
%\section{Acknowledgments}
%We thank our user study participants and the a