### 2014-04-05 doc/uist14/applications.tex
outlined the uist paper

\section{Applications}
\label{sec:applications}

In this section we describe four possible applications that can be built on top of the head orientation-based selection. We focus our discussion on the ``universal remote control'' system.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: