### 2014-04-05 doc/uist14/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper we have presented our system and solutions for the head orientation selection.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: