### 2014-04-05 doc/uist14/disambiguation.tex

\section{Disambiguation Techniques}
\label{sec:disamb-techn}

This section discusses the three different disambiguation techniques we have proposed.

\subsection{Using IR Intensity}
\label{sec:using-ir-intensity}

\subsection{Using Glass Sensors}
\label{sec:using-glass-sensors}

\subsection{Manual Disambiguation}
\label{sec:manu-disamb}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: