### 2014-04-05 doc/uist14/discussion.tex

\section{Discussion}
\label{sec:discussion}

We discuss a few issues with the system.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: