### 2014-04-05 doc/uist14/evaluation.tex

\section{Evaluation}
\label{sec:evaluation}

This section we present our work on the evaluation and user study.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: