### 2014-04-06 doc/uist14/system.tex
refining system section

\section{System Design and Prototyping}
\label{sec:syst-design-prot}

This section describes the design and implementation of our prototype system. We start by summarizing the design considerations in this head-orientation targeting, and for each consideration, we make the choice of commercial off-the-shelf (COTS) components. Afterwards, we conclude this section by a depiction of the overall system architecture that is used for both our study and future application development.

\subsection{Design Considerations}
\label{sec:design-cons}

To effectively enables the physical targeting based on user's head movement, and build such a platform for applications, we have come up with the following design considerations and the justification for each.

{\bf Directional area selection:} To be aligned with the user's head orientation, the sensing modality should be directional. Merely direction is not enough, from Section~\ref{sec:background}, the head muscle's fundamental limitation and Fitts' law imposes us to choose an area selection technique. However,  with a larger coverage area, there will be a higher chance for disambiguation. We should find the proper technique that achieves area selection while still providing ways for a more fine-grained resolution.

% the two visual feedback seems a bit random...
{\bf Leverage visual feedback:} Since the user cannot see his own head orientation, there should be some effective ways to provide feedback to the users. We consider two types of visual feedbacks. The first comes from the environment which enables instantaneous visual cue while the user is looking around. The second should remain unchanged even if the users' head orientation has changed; in such a way, further interactions are enabled without imposing the burden of staring. But such a display shouldn't distract the user's current line-of-sight attention -- which means a near-eye display is better than using a hand-held device.

% A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption. Such feedback should be delivered instantaneously, while users look around a room.

{\bf Flexible communication:} The system is for the interaction with the environment, therefore the communication among devices should be flexible. Especially there would be a consensus problem if the end-nodes are not agreeing on what the user is looking at. The communication would be necessary for resolving the inconsistency in such a distributed environment. Also such communication capability can make the system as a platform for other applications.

{\bf Adjustable angle:} From our pilot study, we have learned that human's head orientation pointing is biased for each person. In order to align area selection technique, there should be ways to adjust or calibrate the covered area.

These design considerations find their expression in our selection of techniques in the following few subsections.

\subsection{IR as beam cursor}
\label{sec:ir-as-beam}

For the directional area selection technique, we have chosen to use infrared (IR) for a couple of reasons.

First, the signal characteristics of IR makes it suitable for directional area selection. There are a wide range of viewing angles that we can choose (DigiKey offers IR emitters from $4^\circ$ to $240^\circ$, however, most are around $20^\circ$ of viewing angle). This leads to a cone shape-like area coverage and can be well-aligned with humna's line of sight.

Second, for the disambiguation consideration, since the infrared illuminance falls off with respect to both the offset angle from the center and the distance to the emitter, even though the signal might cover multiple targets, using the intensity from the receivers can help disambiguation.  
% for the 4 degree to 240 degree, I used information from http://www.digikey.com/product-search/en/optoelectronics/infrared-uv-visible-emitters/524328?k=IR

In addition, IR has been used for remote control for many years. The emitters and receivers are fairly cheap to manufacture; and it's easy for integration with existing devices. 
What's more, IR is immune to most interference sources. The ambient IR might influence the IR receivers' intensity reading, however with simple IR signal validation, the receiver is still safe. There might be issues for signal reflection, either bouncing from the ground to the target or bouncing back to the opposite direction. However, such visual feedback is not aligned with the user's attention so no further action from the user might be taken; or the genuine target could easily beat the false positive since the one with reflection tends to travel a longer distance, making the signal strength smaller.

Therefore, we can attach an IR emitter to capture the head orientation and install some IR receivers in the environment for both signals detection and intensity reading. For our prototype, we have found some COTS components for IR. 
The IR emitter is the one from OSRAM Opto Semiconductors Inc with manufacturer part number SFH 4545. From the datasheet, it has a view angle of $10^\circ$. Part of the reason we have chosen it is because the radiant intensity can quite high if we provide it with sufficient current flow (550mW/sr @ 100mA). We can then easily adjust it using a resistor to satisfy different demands. 
% transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND
For the IR receiver, we use 38.0kHz IR Receiver Modules from Vishay Semiconductor Opto Division (manufacturer part number TSOP38238). 
% receiver webpage http://www.digikey.com/product-detail/en/TSOP38238/751-1227-ND/1681362
In order to read the IR intensity, we integrates another IR light-to-voltage converter from AMS-TAOS USA Inc (manufacturer part number TSL267-LF) which is pretty sensitive to IR irradiance. 
% http://www.digikey.com/product-detail/en/TSL267-LF/TSL267-LF-ND/3095052

We use Arduino platform for the prototyping, and there is an easy-to-use IR library\footnote{\url{https://github.com/shirriff/Arduino-IRremote}} where we can customize the IR signal for both transmission and reception. To read the IR intensity, an 10-bit ADC is attached to the output of TSL267-LF which gives a reading from 0 to 1023.

\subsection{Environment LEDs and Head-up display}
\label{sec:head-up-display}
As we have argued for the visual feedback design considerations, the environment should provide instantaneous visual cue to respond to user's head orientation. This can be easily achieved with 1-bit LEDs on each devices that has an IR receiver. For the visual feedback that is persistent even when the user's head orientation has changed, we have considered a few COTS solutions that support the head-up display, and ended up with Google Glass because it's easy to program (it supports standard Android programming) and the device is also light-weight for everyday use. At the time of writing, we have Google Glass runs on Android 4.0.4 (software version: XE12).

\subsection{Wireless Communication}
\label{sec:wirel-comm}
The IR can be only used for expressing the user's targeting intention, we need another data communication technique that is more reliable and flexible. This is not only necessary for our targeting experiement to transmit received IR signals and the IR intensity readings, but also useful for application development that can exchange arbitrary application-specific messages. We have considered WiFi, Bluetooth Low Energy (BLE) and 802.15.4 (ZigBee) radio. A detailed comparison is out of the scope of this paper, but we have chosen to use 802.15.4 radio, since it is designed for personal area network (PAN) communication, and there are both industrial activity to promote it \footnote{XBee: http://www.digi.com/xbee/} and academia research \cite{watteyne2012openwsn} to make it lower power and more suitable for IoT.

\subsection{3D Printed Adjustable Holder}
\label{sec:3d-print-adjust}
Since we have chosen to use IR for targeting, the holder where IR emitter resides can be designed for adjustment. We have designed a ball-joint model and 3D printed it. See Figure 1.

\subsection{Overall System}
\label{sec:overall-system}
So far, we have described our choice separately, in this subsection we focus on our system architecture that stitches the components together. 
Figure 2 is our overall architecture photo. Because we need to take command from Google Glass, we have used another microcontroller that speaks Bluetooth to communicate with the Google Glass.

%% show the figure here

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: