### 2014-04-12 doc/uist14/relatedwork.tex

%!TEX root = uist14.tex
\section{Background and Related Work}
Our approach is related to head and eye-controlled interfaces, selection techniques, and prior work on hardware devices for pointing in physical spaces.
\subsection{Head and Gaze Input}
Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assitive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, human neck muscles have a much lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
Prior work often focused on used head orientation for control of graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

Gaze can also be used to control graphical user intefaces~\cite{kumar2007eyepoint}. While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map of each space with known target locations. Our system works through point-to-point IR communication and does not require any a priori map. Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}. Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.
%Our work is closer in spirit to Selker's headworn system~\cite{Selker:2001:EGE:634067.634176}.

\subsection{Area Cursors}
A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input. In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. Area cursers are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. We argue that head orientation pointing has similar challenges (limited pointing performance and accuracy). In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed. 
We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement. We contribute a novel physical method based on received light IR intensity.

\subsection{Pointing in physical spaces}
Rukzio studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.

Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. In some techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user what is being selected; however, their small target area make them poorly matched to head orientation input.
Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.

Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar in spirit but relies on simple, single low-cost IR emitters and detectors. 
%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}