### 2014-04-14 doc/uist14/abstract.tex
structural layout as discussed on sunday - iterative development

%!TEX root = uist14.tex
Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
%
We introduce and characterize a method for selecting and interacting with physical targets at a distance using head orientation. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. The fundamental imprecision of head orientation requires disambiguation techniques when the intended target is unclear. To improve on the slow manual disambiguation of our first design, we introduce a second design in which receivers use IR intensity sensors to estimate which target is the intended selection. A second study demonstrates that using IR intensity successfully reduces ambiguity and enhances selection speed. Finally, we describe how a spatial data structure can be inferred to enable a natural mapping of head direction during the disambiguation step and we report first use data. 
We demonstrate an application of our selection technique for controlling smart appliances and report qualitative user impressions.