### 2014-04-14 doc/uist14/abstract.tex
a significant amount of new content and changes on papers

%!TEX root = uist14.tex

\ben{This parts needs to be changed based on the current flow of the story.}
Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
%
We introduce and characterize a method for selecting and interacting with physical targets at a distance using head orientation. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. We present our iterative design process that has shown with IR, it outperforms a pure list selection. And with IR intensity, the time can be further reduced. If head motion is added, we can infer a spatial data structure of the nodes to enable a natural head-motion based disambiguation.
We demonstrate an application of our selection technique for controlling smart appliances and report qualitative user impressions.