### 2014-04-14 doc/uist14/discussion.tex

\section{Discussion}
\label{sec:discussion}

This section is about a few issues in our paper.

\subsection{Using Google Glass}
\label{sec:using-google-glass}
From one of the reviews during BiD paper party, people are suggesting a discussion what we get and what we lost of using Google Glass.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: