### 2014-04-14 doc/uist14/iteration1.tex

%!TEX root = uist14.tex

\section{Iteration 1: IR vs. List}
Our system design uses IR as the beam cursor for physical target selection because the area coarse selection will remove the targets outside of users' view range. And this could help during the fine selection time because there is a smaller set of targets for fine selection. The opposite is to have a complete list of all targets when there is no single coarse selection and the disambiguation is always needed. This brought up the first question that we asked ourselves -- {\em how much does using IR help in reducing the selection time}. To answer that question and quantify the benefit of having IR for the coarse selection, we carried out our first target acquisition study which compares the {\em IR selection} and {\em list selection}.

We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}). Each of them has a number as ID and a letter representing the name. There are 14 participants from our institution for this study. 13 had never used Glass before, therefore we offered a Google Glass crash course before the experiment for them to get familiar with Glass gestures. 4 of them wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment. Targets were called out by number; for the list mode condition, participants need to match numbers to letters.}
\label{fig:targeting-study-layout}
\end{figure}

In the within-subject study, half of them performed {\em IR selection} first and the other half did {\em list selection} first. For each selection condition, we conducted 15 target selection by random choosing from all the targets. During the study, we measure the {\bf target acquisition time} for each target selection, and afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.

%% remove this for now
%% Our measurements suggest that IR communication can be targeted to an area about 60-120cm in diameter, up to 5m in front of the user. These values are a reasonable match for selecting targets in a room-size environment. A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously. A narrower beam will make targeting more challenging, given the precision constraints of human head movement.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B). The centers of boxes are median values, while white dashed lines are mean values.}
\label{fig:selection-time}
\end{figure}

Our results indicate that {\em IR selection} is generally better than {\em list selection}. From Figure~\ref{fig:selection-time}A, the average target acquisition time for {\em IR selection} is 6.67 seconds, while for {\em list selection} it took 8.86 seconds. We run Student's t-test on the data and it shows significant difference ($t(279)=-3.81, p=0.00017$).

To further understand if the gain comes from the coarse selection, we splited the data from {\em IR selection} to two parts -- ones that require disambiguation and ones that don't. The result is shown in Figure~\ref{fig:selection-time}A. It shows that it takes 6.40 seconds (on average) to complete the connection if coarse selection has achieved the acquisition, while when multiple devices are in range, the time increases to 9.16 seconds, indicating additional 2.76 seconds for disambiguation. Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the latter case, they only happen 10\% of total {\em IR selection} trials.

From the above result, we learned that with IR, we can have effectively reduced the overall target acquisition time. And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
\caption{Times taken to select a device vs.~its order in the list. The dotted line is a linear fit between the mean times and device orders. Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
 % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
\label{fig:time-vs-list-order}
\end{figure}

To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in {\em list selection} condition, ordered by their relative position in the list. Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}. From this figure, we can see that once the target's order is larger than 6, the average acquistion time will be larger than the case when disambiguation is needed in {\em IR selection}. Therefore, we argue that once there are more targets in the physical space, the {\em list selection} is not scalable.

\ben{qualitative feedback hasn't been added yet for this part.}
%\subsubsection{Preference}

%Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome. As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.

%List mode had certain advantages: It was judged to be more accurate and predictable as there was always exactly one device selected in the list (\studyquote{With the list you never have to worry about accidentally picking up two targets}). Also, it did not require a clear line of sight to the target device so participants did not have to move from their starting position (\studyquote{The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance}).

%On the other hand, list mode was judged to be more \studyquote{annoying} and tedious. The temple-based touchpad for selection was difficult to use for a participant with long hair: \studyquote{List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.} Another participant also commented on the ergonomic challenge of touchpad use on Glass: \studyquote{The strength of the IR mode was that I didn't have to use my fingers as much to control. If the items were spaced relatively far apart, it was easy to select a specific appliance.}

%One noted benefit of infrared mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen. One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.

%A few perceived weaknesses of infrared mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target. One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}. Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}. Users had to learn the usable angle of the IR emitter before they became successful at controlling the devices: \studyquote{I had to compensate by tilting my head up a little bit.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: