### 2014-04-16 doc/uist14/abstract.tex
updated a bunch of figures

%!TEX root = uist14.tex

%% Yang suggests: Interacting with smart objects in the physical space efficiently is a realistic challenge as these objects become ubiquitous. In this paper, we contribute HOBS, a set of novel methods for selecting physical objects at a distance using infrared-sensed head orientations. We augment a commercial head-worn device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. We present the iterative design process of our methods, involving a series of interaction technique and hardware design and user evaluations....
\ben{We should remove hands-free?? which is orthogonal to our contribution.}

Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
%
We present the iterative design and evaluation of \systemname -- a head-orientation based selections method for such devices, for interacting with physical targets at a distance. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. Our first design shows that a naive IR implementation can outperform list selection, but has poor performance when refinement between multiple targets is needed. A second design uses IR intensity measurement at targets to improve refinement. To address a lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets from demonstration to enable a natural head-motion based disambiguation.
%
We also demonstrate a universal remote control application of the technique and report qualitative user impressions.