### 2014-04-16 doc/uist14/conclusion.tex
applications, conclusion, discussion, iteration3: grammar/minor content

\section{Conclusion}
\label{sec:conclusion}

In this paper, by presenting our iterative design process in head-orientation based targeting, we have learned that IR alone can help reduce the overall acquisition time by reducing the chances when we need to perform refinement. With IR intensity added, the targeting can work better in a relative dense environment. However, a more natural approach is to combine IR with head motion. Through our preliminary user studies, we learned that this is a more intuitative way of performing refinement in comparison to the menu-based selections. We leave a more comprehensive technical solution of using motion sensors and its evaluation as the future work.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End: