### 2014-06-23 doc/sui14/sui14.tex

\documentclass{sigchi}

% Use this command to override the default ACM copyright statement (e.g. for preprints). 
% Consult the conference website for the camera-ready copyright statement.

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is 	granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org. \\\textbf{}
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
\toappear{Submitted for review to SUI 2014.}

% Arabic page numbers for submission. 
% Remove this line to eliminate page numbers for the camera ready copy
\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs
% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

\usepackage{gensymb}
\usepackage{afterpage}
% see http://tex.stackexchange.com/questions/46055/typesetting-with-inch-symbols-and-sizes-in-inches
% \usepackage{mathpazo}
\usepackage{amsmath}
\def\inch#1{#1''}
\def\ft#1{#1'\thinspace}

% comment macros are in inputmacros.tex
\input{inputmacros}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={SIGCHI Conference Proceedings Format},
pdfauthor={LaTeX},
pdfkeywords={SIGCHI, proceedings, archival format},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

% as we discussed last time, the "Sees" and "Line-of-sight" suggests some computer vision direction 
% need a different title for sure, but not in a hurry
\title{\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces}
%Project GlasSees: Direct Interaction Through Line-of-sight}

\iftoggle{anonymous}{
\author{
 \alignauthor Anonymous for submission\\
    %\affaddr{...}\\
    %\email{...}\\
  }
}{ %else
  \numberofauthors{1}
  \author{
  \alignauthor  Ben Zhang$^{\dagger}$, Yu-Hsiang Chen$^{\dagger}$, Claire Tuna$^{\dagger}$, Achal Dave$^{\dagger}$,\\Yang Li$^{\ddagger}$, Edward Lee$^{\dagger}$, Bj\"orn Hartmann$^{\dagger}$ \\
  \affaddr{$\dagger$: UC Berkeley EECS \& CITRIS Invention Lab \hspace{0.25in} $\ddagger$: Google Research}\\
  \email{\{benzh,clairetuna,achal,eal,bjoern\}@berkeley.edu, sean.yhc@gmail.com, yangli@acm.org}
  }
}

\teaser{
  \centering
  \includegraphics[width=\textwidth]{figures/teaser.png}
  \caption{{\em Left}: Our head orientation-based selection techniques use an IR emitter -- multiple targets may fall within its illumination area. {\em Center}: We offer two list-based refinement techniques -- {\em Naive IR} uses alphabetical ordering; {\em Intensity IR} orders targets by IR intensity. {\em Right}: Alternatively, users can refine their selection through head orientation refinement in a quasi-mode.}
  \label{fig:teaser}
}

%% the third part of the teaser: An example application of head-orientation targeting: Using an augmented head-worn device (1), users can control smart home appliances (2) by looking at them. A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}
\vspace{-0.05in}
\keywords{
  Wearable computing; selection; glass; infrared
}
\vspace{-0.1in}
%% refer to http://www.acm.org/about/class/ccs98-html
\category{H.5.2.}{Information Interfaces and Presentation (e.g. HCI)}{Interaction styles (e.g., commands, menus, forms, direct manipulation).}

\input{intro}
\input{relatedwork}
\input{interaction}
\input{iteration1}
\input{iteration2}
\input{iteration3}
%\input{disambiguation}
%\input{evaluation}

%% the rest of the paper seems to be fine; it's mainly the texts
\input{applications}
\input{discussion}
\input{conclusion}

\iftoggle{anonymous}{
% no acks in anonymous submission
}{
  \section{Acknowledgments}

  This work was supported in part by the TerraSwarm Research Center, one of six centers supported by the STARnet phase of the Focus Center Research Program (FCRP) a Semiconductor Research Corporation program sponsored by MARCO and DARPA. Additional support was provided by a Sloan Foundation Fellowship and a Google Research Award.
}
%figure template:
%\begin{figure}[!h]
%\centering
%\includegraphics[width=1.0\columnwidth]{Figure1}
%\caption{With Caption Below, be sure to have a good resolution image
%  (see item D within the preparation instructions).}
%\label{fig:figure1}
%\end{figure}

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
%% \balance

\bibliographystyle{acm-sigchi}
\bibliography{sui14}
\end{document}