### 2014-06-27 doc/sui14/abstract.tex
upper case acronym letters in abstract, minor grammatical fixes

%!TEX root = uist14.tex

%% Yang suggests: Interacting with smart objects in the physical space efficiently is a realistic challenge as these objects become ubiquitous. In this paper, we contribute HOBS, a set of novel methods for selecting physical objects at a distance using infrared-sensed head orientations. We augment a commercial head-worn device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. We present the iterative design process of our methods, involving a series of interaction technique and hardware design and user evaluations....
Emerging head-worn computing devices can enable interactions with smart objects in physical spaces.
%
We present the iterative design and evaluation of \systemname -- a Head-Orientation Based Method for interacting with these devices at a distance. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. Our first design shows that a naive IR implementation can outperform list selection, but has poor performance when refinement between multiple targets is needed. A second design uses IR intensity measurement at targets to improve refinement. To address the lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets enabling a natural head-motion based disambiguation.
%
Finally, we demonstrate a universal remote control application using HOBS and report qualitative user impressions.