### 2014-06-27 doc/uist14/submission/rebuttal.txt
edits

We thank the reviewers for their detailed feedback. Below, we answer the
raised questions and clarify some points of confusion.

CONTRIBUTION  LEVEL (AC, R1, R2, R3)
Reviewers find this work’s contribution limited and suggest focusing on
interacting with complex targets. In the past, we built interfaces to
control complex appliances including smart TVs; however, CHI reviewers on
a prior application-centric paper suggested focusing on fundamental
selection techniques instead, resulting in the abbreviated description of
applications in our paper. Some examples can be found in the APPLICATION
section; we are happy to reintroduce more material in our revision.

We focus on the fundamental question of target selection itself, which is
likely to be a problem in a future rich with smart devices. Existing
commercial solutions (primarily list selection) don’t scale well (we
envision a swarm of interactive devices as the contexts) and exhibit
naming problems when multiple identical objects are within the same
space. Our work contributes a sophisticated understanding of methods for
using head-orientation as an indication of user’s point of interest in
physical space.

LOW DISAMBIGUATION SUCCESS RATE (R1)
R1 comments that our technique barely outperforms chance. However, our
success rate of 55% in iteration 2 is much higher than random chance
given that there are at least 4 targets in the user’s field of view in
our setup (in this case, chance would be 25%).

COMPUTER VISION APPROACH (AC, R2, R3)
Reviewers suggest considering computer vision-based techniques with
markers in the environment (such as QR codes, or Bokode [Mohan] which
extends the working distance to a few meters for visual tags). We
considered this direction and found several drawbacks, including
computational complexity, low accuracy when targets are far, the absence
of real-time environment feedback with passive tags, and aesthetic issues
with QR codes. While we don’t claim our choice of IR is optimal, it has
important advantages -- it is low-cost, readily available and quite
suitable for area selection. We can weaken our claim and add explanations
of how disambiguation techniques might generalize to other methods of
implementing target identification.

SYSTEM DESCRIPTION (AC)
The AC commented that “the description of the system is very
lacking”. We provided brief implementation information in order to
focus on the design and evaluation within a shorter paper to match the
magnitude of contribution. This may have been counterproductive. We list
part numbers in the "Hardware" subsection but can describe how IDs are
exchanged over IR and 802.15.4 in more detail.

MISSING RELATED WORK (AC, R3)
We will add the additional references.
RFIG Lamps and photosensing wireless tags [Raskar et al] use a projector
and data is encoded in each pixel so that tags can localize themselves
for target selection. Our IR emitter, a relatively low-cost solution,
doesn't have this feature but shares the commonality of covering an area
for the ease of selection. At the end of their paper, they envision an
IR-based system to solve ambient light problems. Our work lies in that
direction and also contributes evaluations with user studies.
Progressive Refinement [Bacima] studies several progressive refinement
selection modalities, which is close to our two-stage selection; but our
contexts are head-orientation as it reflects users’ point of interest.
Tinmith-metro [Piekarski et al] tracks the users’ hand movements and
uses a HUD for augmented reality, while our work focuses on direct
selection on physical targets.

MISC (AC)
The AC asked why users couldn't swipe in a certain direction on Glass for
disambiguation. Two major factors preclude this interaction: 1) the Glass
physical form factor makes upward swipes awkward to perform; 2) the plane
of the touchpad is orthogonal to the plane of the targets, leading to
ambiguity what lateral swipes mean.