### 2014-06-28 doc/sui14/iteration3.tex

%!TEX root = uist14.tex
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}

Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the
near-eye display for the refinement step. The list interface has a few clear
shortcomings: navigation actions map poorly to real world results, and
users must switch their focus back and forth between the physical scene and the
near-eye display. These problems motivated us to design an approach that
harnesses the head orientation during the {\em refinement} stage.

In our third iteration, we ask the question {\em can we build a system purely
using head orientation and visual feedback from the environment for target
selection?}

\subsection{Technique}

We introduce a third technique that uses a combination of motion sensors and IR
to learn the relative orientations of targets in a room and intelligently
suggest targets during refinement (see Figure~\ref{fig:third_technique}).

In the learning stage, the user scans over the targets, and the system attains
the absolute orientation of each device from IR and motion sensors. From this
information it can abstract out their relative positions and build an {\em
adjacency map}.

This absolute orientation cannot be applied to all indoor environments, since
the user’s movements through the space could change the relationships between
targets. However, with the constraint that the targets are spread around the
periphery, their relative orientations are stable (see
Figure~\ref{fig:third_principle}).

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/third_principle.png}
\caption{This illustrates that the change of a user's absolution position
doesn't change the relative relationship of physical targets.}
\label{fig:third_principle}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/third_technique.pdf}
\caption{Our third technique learned each target's absolute orientation and
construct the adjacency map. During the {\em refinement} stage, the prediction
is based on relative changes to a reference point.}
\label{fig:third_technique}
\end{figure}

After the map is created, the user can hold down on the touchpad to enter a
quasi-mode for refinement. In this quasi-mode, one device lights up at a time.
When the user turns their head in the direction of another device, the light
switches to that device. Therefore, the user can move between devices one at a
time with slight head movements. We implement this interaction by calculating
the user's direction of motion using a low-passed history of sensor
measurements and searching through the adjacency map for the nearest device in
that direction.

\subsection{Evaluation}
We evaluate the head motion refinement method by holding an informal study and
collecting qualitative feedback from a subset of 4 users from the Iteration 2
study. In the study, we asked users to try cycling through the targets using
the new quasi-mode.

The users had strong preferences for the new method of refinement. On a scale
from 1-7, 1 being the least mental effort and 7 being the most mental effort
users rated the old technique 4.25 and the new technique 2 on average. 100\%
indicated a preference for head movement to list navigation. One user
referenced the issue of naming targets that the list necessitates, preferring
the experience of \studyquote{matching visual cues rather than numbers}.
Another participant remarked that it \studyquote{just made more sense} and was
a \studyquote{more natural way for demonstrating intentionality}. The users
preferred the new mapping in relation to the whole environment: \studyquote{it
leveraged the spatial sense that I already had just by using the system}. They
were also delighted to avoid list navigation, which they now called
\studyquote{difficult} and \studyquote{painful}.