### 2014-08-11 doc/sui14/iteration3.tex
refining texts for iteration3, more descriptions are needed

%!TEX root = uist14.tex
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}

Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the
near-eye display for the refinement step. The list interface has a few clear
shortcomings: navigation actions map poorly to real world results, and
users must switch their focus back and forth between the physical scene and the
near-eye display. These problems motivated us to design an approach that
harnesses the head orientation during the {\em refinement} stage.

In our third iteration, we ask the \changes{research} question --
{\em can we build a system purely using head orientation and visual
  feedback from the environment for target selection?}

\subsection{Technique}

We introduce a third technique that uses a combination of motion sensors and IR
to learn the relative orientations of targets in a room and intelligently
suggest targets during refinement (see Figure~\ref{fig:third_technique}).

In the learning stage, as the user scans over targets in their normal
use, the system attains the absolute orientation of each device from
IR and motion sensors. From this information it can abstract out
\changes{the} relative positions \changes{of targets} and construct an
{\em adjacency map}. The absolute orientations in the map cannot be
applied to all indoor environments, since the user's movements through
the space could change the relationships between targets. However,
with the constraint that the targets are spread around the periphery
\changes{at similar distances}, their relative orientations are stable
(see Figure~\ref{fig:third_principle}). And since we are using IR at
the same time, the learning process is transparent to the user and the
map can be built without an explicit calibration phase.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/third_technique.pdf}
\caption{Our third technique learned each target's absolute orientation and
construct the adjacency map. During the {\em refinement} stage, the prediction
is based on relative changes to a reference point.}
\label{fig:third_technique}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/third_principle.png}
\caption{This illustrates that the change of a user's absolution position
doesn't change the relative relationship of physical targets.}
\label{fig:third_principle}
\end{figure}

\changes{After the map is created, the user can enter a quasi-mode for
  refinement by holding down on the touchpad. In this quasi-mode, one
  device lights up at a time. When the user turns their head in the
  direction of another device, the light switches to that
  device. Therefore, the user can move between devices one at a time
  with slight head movements. This prediction is implemented by
  calculating the user's direction of motion using a low-passed
  history of sensor measurements and searching through the adjacency
  map for the nearest device in that direction.}

\subsection{Evaluation}
We evaluate the head motion refinement method by holding an informal study and
collecting qualitative feedback from a subset of 4 users from the Iteration 2
study. In the study, we asked users to try cycling through the targets using
the new quasi-mode.

The users had strong preferences for the new method of refinement. On a scale
from 1-7, 1 being the least mental effort and 7 being the most mental effort
users rated the old technique 4.25 and the new technique 2 on average. 100\%
indicated a preference for head movement to list navigation. One user
referenced the issue of naming targets that the list necessitates, preferring
the experience of \studyquote{matching visual cues rather than numbers}.
Another participant remarked that it \studyquote{just made more sense} and was
a \studyquote{more natural way for demonstrating intentionality}. The users
preferred the new mapping in relation to the whole environment: \studyquote{it
leveraged the spatial sense that I already had just by using the system}. They
were also delighted to avoid list navigation, which they now called
\studyquote{difficult} and \studyquote{painful}.